### 2011-10-04 doc/manuscript.tex
It is not clear how the manuscript is made.

\section*{Abstract}
Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  We aims at which parts of the genomic
organization, including human pathogenic virulence factors, of SDE could have
originated in those of SPY, or vice versa.  Our genome-wide survey of gene
transfer indicated that gene transfer via homologous recombination occurred
relatively from SPY to SDE.  Although we did not find enrichment of virulence
genes in the homologous recombining gene transfer, we found evidence that they
transferred more often via non-recombining fashion without replacing parts of
destination genomes than did non-virulence genes.  A genome-wide resource of
recombining and non-recombining gene transfer for the study of the genomes is
provided so that research community in pathogenic bacteria could take advantage
of it.

@@

Therefore, the research topic on gene exchanges between SPY and SDE has been
discussed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyrogenic
exotoxins \textit{speG}. \citet{Davies2007} investigated phage 3396 from SDE as
a vehicle of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that
an integrative conjugative element of SDE allowed gene transfer from SDE to
other streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
to SDE was more dominant than the other direction.  Because these researches on
the controversial issue, directionality of gene transfer between SPY and SDE,
have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing database, it remains to
be elusive to determine the genome-wide degrees to which genomic segments flow
between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal
of genomes available.  Although this type of study gave us insights into complex
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
investigated gene transfer events in \textit{Eschericha Coli} using the embedded
quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
\citet{Hamady2006} combined phylogenetic and compositional approaches to
detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
of \textit{Shewanella baltica} genomes to identify the genetic elements that
enabled the species to adapt to redox gradients.  Although several complete
genomes from SPY have been available, \citet{Suzuki2011} recently added one more
complete genome of SDE to the previous set of a unique SDE genome. This set of
SPY and SDE genomes prompted us to study genome-wide gene flows between the two
species that afflicted human beings around the globe.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within
the same lineage.  Mutations on a gene modifiy relatively small portions of the
gene through substitutions, insertions, and deletions of deoxyribonucleic
acids. Horizontal gene transfers can reshape the organization of a bacterial
genome, which could revamp biological functions of the bacterial species.
Recipient bacteria of transferred genes can substitute transferred genes for homologous parts of
their genomes, or integrate transferred genes into their genomes
without replacement. These two processes are here in this paper discriminated by calling the
former \textit{recombining gene transfer}, and the latter \textit{horizontal
gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
evolutionary processes in bacteria, reseraches on statistical analyses of
complete genome sequences in efforts of better understanding bacterial evolution
have paid less attention to a combined approach to illuminating disparity of
the two gene transfer processes. We attempted here to use and compare two
approaches for distinguishing recombining gene transfer process from horizontal, or
non-recombining, gene transfer one, and more systematically tackled their roles
in association with virulence genes.

Homologous recombining gene transfer could be studied more readily using the
core genomes from multiple individual bacterial genomes while horizontal gene
transfer might be more easily detected using the pan genomes. Recently, a method
of inferring homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacterial individuals of \textit{Bacillus
cereus} species in the core genomes.  More complex evolutionary processes acting
on bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  Because of the complexity, these processes do not
readily lend themselves to statistically more sound model-based approaches such
as maximum likelihood. Instead, a parsimony-based approach to inferring gene
gain, loss, and transfer events was developed by  \citep{Doyon2011}.  We noted
that the first approach could be used to infer transfer of genomic segments due
to homologous recombination, and the second could do similarly attributed to
horizontal gene transfer. Although this heterogeneously combined approach could
be unsatisfactory, we hoped to take the unique opportunity of new genomes to
study gene transfer in genome-wide scale among SPY and SDE individual bacteria.

In short, the complete genomes of the two spcies of SPY and SDE presented a unique
opportunity to study gene transfer over relatively recent time scales.  We used
both of the model-based approach for inferring homologous recombination and the
parsimony-based one for detecting gene gain, loss, and horizontal transfer. We
also compared both of the results in recombining gene transfer events to show
that the parimony-based method inferred gene transfer events in similar
biological functional categories to those inferred with the model-based
approach. This is the first attempt as far as we know to analyaze SPY and SDE in
genome-wide scale for detecting gene transfer events not only in the core genome
but also in the pan genome, which make the analysis complete in terms of the
amount of used genome data. Our analysis of core genomes also showed for the
first time that the net directionality of, at least, recombining gene transfer
from SPY to SDE was stronger than the reverse direction.

@@

\subsection{Recombination-driven substitutions were stronger than mutation-driven ones}

The genome alignment of the five genomes (Table S\ref{tab:genome}) showed two distinct
genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE genomes were
larger in length than those of SPY. Although it was unclear that SDD was
different from either of SDE1 and SDE2 from the graphical view of the genome
alignment, the spcies tree inferred from the genome alignment showed that SDD
branch was attached to the clade branch of SDE1 and SDE2 (Figure
S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
implied that the model assumption of geometric length distribution might be
plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
Method} and \citealt{Didelot2010}).  The relative strength of recombination
compared to mutations indicated that reconmination-driven substitutions affected
more sites than did mutation-driven substitutions (Table
S\ref{tab:clonalframe}). In the \textit{Streptococcus} bacteria homologous
recombination appeared to be a stronger evolutionary force than point mutations 
process.

@@

Along the genome of SDE1 population mutation rate did not appear to change much
(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
mutation rates were larger than twice the global estimate of mutation rate.
% \texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
The recombination rate appeared to change more dynamically. Espeically, the
genomic region of SDE1 near 20 kilobase pairs from the origin of replication
contains alignment blocks with bell-shaped recombination rates.
% \texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}
The recombinant tract length changes in a way that we could not note any
particular positions along the genome.
There appeared to be negative correlation between recombinant tract length and
recombination rate.
% \texttt{PLOT PAIRWISE SCATTER PLOTS OF THE THREE PARAMETERS.}

@@

Most of local tree topologies (see section \textbf{Materials and Methods}), 
or 66.7\%, along the alignment blocks were the
same as the species tree, which allowed us to be more confident of the species
tree inferred using the alignment blocks (Table
S\ref{tab:Gene-tree-topologies}).  The second most probable topology (9.3\%)  was one
where SDD was the outgroup, and SPY and SDE were sister clades 
SPY and SDE appeared to share about tenth of the core
genomes that they were more closely related with each other than each to SDD's
genome. The other topologies had less than 4\% of proportions.

@@

The heat map in Figure \ref{fig:Heatmap-of-recombination} shows the log base 2
ratios of the \textit{a posteriori} number of recombinant edges for a pair of
donor and recipient species tree branches and their corresponding expected
values given the species tree and the global parameter settings.  Because
recombinant edges from any species tree branch to the root branch are not
possible under \texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}),
the column for the root as recipient was colored green. Similarly, all of the
cells where a recipient branch was older than the donor, and two branches did
not share any contemporary time were colored green. The diagonal cells were
colored gray, indicating no significant number of recombinant edges, because the
observed number of recombinant edges and the expected one were similar. As
recombinant edges starting from and ending in the same species tree branch would
not change the species tree, which would not change the likelihood much, the
cells should be colored gray.  Because SDD is a strictly veterinary pathogenic
bacteria, we did not expect that SDD received much gene transfer from either
SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was -2.20, and that from
SPY2 to SDD was -1.97. On the contrary, the recombing gene transfer from SDE1 to
SDD was 0.756, and that from SDE2 to SDD was 0.698. This relatively higher
recombining gene transfer from two SDE individuals to SDD compared with that
from two SPY individuals to SDD reflected that the SDD was more closely related
to SDE than to SPY in most of the genomes.

The genomic organization between SDD and SDE1/SDE2 were much similar (Figures
S\ref{fig:mauve}), and they formed a clade in the species tree
(S\ref{fig:tree5}). However, the gene transfer from SDD to either SDE1 or SDE2
was less frequent than the case of reverse direction; SDD served as a recipient
rather than a donor. It is not unclear why SDD received more genes from SDE1 or
SDE2 than did SDE1 or SDE2 from SDD.

The recombining gene transfer log base 2 ratios between SDE1 and SDE2 (1.17 and
1.20) were about twice as large as that from SDE1 and SDE2 to SPY1 and SPY2
(0.67, 0.66, 0.60, and 0.39).  The intensities of recombining gene transfer from
external branches of SDE to external branches to SPY was similar to those from
SDE1 or SDE2 to SDD (0.76 and 0.70).  Either SPY1 or SPY2, and SDD appeared to
be no different from the view point of SDE1 or SDE2 because of the similar
values of recombining gene transfer intensity from SDE1 or SDE2 to the other
three external branches.  It was expected that the recombining gene transfer
within a clade would be larger than that between clades.  Surprisingly, the
recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
1.07) was comparable with those from SPY1 and SPY2 to SDE1 and SDE2 (1.25, 1.25,
1.17, and 1.29).  Effectively, the recombining gene transfer from a clade SPY1
or SPY2 to a different clade SDE1 or SDE2 was no different from that within the
clade of SPY1 and SPY2. The two external branches of SDE received genes from
either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2 were in
the same clade as SPY1 and SPY2.  The bacterial recombination between SPY and
SDE might happen more often than that between SDE and SDD whereas the genome
configurations of SDE and SDD are more similar than each to SPY. This might be
consistent with the fact that SPY and SDE shared the same ecological niche.  The
gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was larger than
any type of gene transfer: 1.88 and 1.94, respectively.  Furthermore, we found a
total of 58 genes for which a recombinant edge of certain types were sampled
with probability at least 0.9 by focusing on specific types of recombinant
edges.  (Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
transferred from SPY to SDE rather than the reverse direction overall.  All
things considered, this clearly showed that the recombining gene transfer was
biased toward the net direction from SPY to SDE.

Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining gene
transfer events between the different branches of the species tree.  SDE1 and
SDE2 shared the most events (26 and 34) compared to other sample pairs.  SDD
appeared to acquire many genes from the ancestral SDE branch. Transfer events
between the SPY and SDE clades were inferred in smaller numbers (<6) compared to
those within each clade (23--34). Among the clade-crossing transfers, we did not
see any significant enrichment for a specific direction, as opposed to the
result for recombining transfers from the model-based analysis (Figure
S\ref{fig:mowgli-recomb-heatmap}).

\subsection{More frequent recombining gene transfer than horizontal gene transfer}
Among the 2314 gene families
the largest family consisted of 59 genes, while
nearly half of gene families (1066) had exactly one gene per species
, or six genes total (Figure S\ref{fig:famsize}).
Figure \ref{fig:Gene-duplication-loss} shows the
inferred gene counts for each extant and ancestral species. Overall,
we inferred 304 duplications, 944 recombining gene transfers, and 294
non-recombining gene transfers, comprising 23.7\% of total transfer
events.  We found that duplications made up only a small fraction of
the events within the clade (10.2\%), while horizontal transfers and
losses were significantly more frequent (41.7\% and 48.0\%,
respectively). We found that duplications were most often
species-specific. Many gene losses likely represented horizontal
transfers that replaced the native gene sequence.  The species SDD and
SDE were the most expanded, both in terms of duplications and in-bound
transfers.

@@

\subsection{Horizontal gene transfer is associated with virulence genes}
\texttt{Select genes with high horizontal gene transfers. Relate these to
categories of virulence genes. I should find many more virulence genes in Matt's
analysis.}

\subsection{Comparison between model-based and parsimony-based approaches}
\texttt{I might want to analyze only core genomes and pan genomes
separately. By separating data set of 6-gene families from the other families,
we could have more complete comparison.}

@@

\section{Discussions}
Closely related bacterial species were investigated here to better
understand gene transfer processes. Two complementary methods were employed to
study recombining and non-recombining gene transfers. By discriminating the two
processes in analyzing the genomes, we provided evidence that the pathogenicity
of SDE and SPY was more attributed to non-recombining gene transfer process than
homologous recombining gene transfer.  Direction of gene flow between SDE and
SPY has been questioned and challenged
\citep{Kalia2001,Kalia2004,Towers2004,Bessen2005,Davies2007,Bessen2010} in order
to elucidate the origin of opportunistic human pathogenicity of SDE via transfer
from SPY into SDE.  We also showed in this study that the amount of recombining
gene flows from SPY to SDE was relatively greater than that of the reverse
direction using the core genome sequences.

\textbf{Virulence genes and gene transfer:} 
Because we found directional recombining gene transfer from SPY to SDE, and
did not found virulence genes association with recombining gene transfer, 
this indicated that 
pathogenicity of SDE might not be due to acquisition of virulence genes from
SPY via recombining transfer. This does not necessarily mean that virulence
genes were not exchanged between SPY and SDE via non-recombining transfer. 
Because the Mowgli analysis just did not reveal unbalanced non-recombining or
recombinding gene transfer, we suspected that
Mowgli's parsimony might not be powerful in detecting unbalanced gene flow.
We conjecture the following hypothesis of the evolution of SPY and SDE to
support the non-recombining virulence gene exchange between SPY and SDE.
Pathogenicity of SDE might have existed before the common ancestor of SPY
and SDE.  We speculated that SPY might have been split off from the common
ancestor that was SDE-like because the population size of SDE was larger than
that of SPY (CITE). After the split, SPY might have been specialized in
inhabting human hosts, and SDE in colonizing veterinary hosts. Recent surge in
cases of human diseases attributed to SDE might be because of secondary
contact of SPY and SDE since their split from their common ancestor.  When SPY
thrives in its environment of human hosts and another relatively close species,
SDE, starts to cohabit with the indigenous species SPY, genetic elements in one
of the two species could benefit another species. If the two species do not
compete with each other, then the increased genetic variation from multiple
species would eventually bring advantage to both of the species. 
Virulence genes could have been beneficial to both species in inhabting human hosts. 
Because the common ancestry of SPY and SDE, recombining gene transfer between
the two species might be relatively prevalent. But, these homologous
recombination might have rarely contributed to influx of virulence genes from
SPY to the opportunistic human pathogen of SDE. Because homologous parts
of SPY and SDE would be similar in their genetic compositions, the homologous
parts would be unlikely to contain virulence genes to human hosts. We suspected
that virulence genes were exchanged between SPY and SDE, but did not find any
directionality of the gene flow.

@@

\section*{Supplementary Information}

>>>>

\section*{Abstract}
Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  We aims at which parts of the genomic
organization and human pathogenic virulence factors of SDE could have
originated in those of SPY, or vice versa.  Our genome-wide survey of gene
transfer indicated that gene transfer via homologous recombination occurred
relatively from SPY to SDE.  Although we did not find enrichment of virulence
genes in the homologous recombining gene transfer, we found evidence that
virulence genes transferred more often via non-recombining fashion without
replacing parts of destination genomes than did non-virulence genes.  A
genome-wide resource of recombining and non-recombining gene transfer for the
study of the genomes is provided so that research community in pathogenic
bacteria could take advantage of it.

@@

Therefore, the research topic on gene exchanges between SPY and SDE has been
discussed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyrogenic
exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as
a vehicle of gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that
an integrative conjugative element of SDE allowed gene transfer from SDE to
other streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the direction of gene transfer from SPY
to SDE was more dominant than the other direction.  Because these researches on
the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal
of genomes available.  Although this type of study gave us insights into complex
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach developed by \citet{Karlin2001}. \citet{Luo2011}
investigated gene transfer events in \textit{Eschericha Coli} using the embedded
quartet decomposition analysis developed by \citet{Zhaxybayeva2006}.
\citet{Hamady2006} combined phylogenetic and compositional approaches to
detecting horizontal gene transfer.  \citet{Caro-Quintero2011} studied a number
of \textit{Shewanella baltica} genomes to identify the genetic elements that
enabled the species to adapt to redox gradients.  Although several complete
genomes from SPY have been available, \citet{Suzuki2011} recently added one more
complete genome of SDE to the previous set of a unique SDE genome
\citep{Shimomura2011}. The multiple SDE genomes prompted us to study
genome-wide gene flows between the two species that afflicted human beings
around the globe.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within the
same lineage.  Mutations on a gene modifiy relatively small portions of the gene
through substitutions, insertions, and deletions of deoxyribonucleic acids.
Horizontal gene transfers can reshape the organization of a bacterial genome,
which could revamp biological functions of the bacterial species.  Recipient
bacteria of transferred genes can substitute transferred genes for homologous
parts of their genomes, or integrate transferred genes into their genomes
without replacement. These two processes are here in this paper discriminated by
calling the former \textit{recombining gene transfer}, and the latter
\textit{horizontal gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the
two distinct evolutionary processes in bacteria, analyses of complete genome
sequences in efforts of better understanding bacterial evolution have paid less
attention to a combined approach to illuminating disparity of the two gene
transfer processes. We attempted here to use and compare two approaches for
distinguishing recombining gene transfer process from horizontal, or
non-recombining, gene transfer, and more systematically tackled their roles
in association with virulence genes.

Homologous recombining gene transfer could be studied more readily using the
core genomes from multiple individual bacterial genomes while horizontal gene
transfer might be more easily detected using the pan genomes. Recently, a method
of inferring homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacterial individuals of \textit{Bacillus
cereus} species in the core genomes.  More complex evolutionary processes acting
on bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  Because of the complexity, these processes do not
readily lend themselves to statistically more sound model-based approaches such
as maximum likelihood. Instead, a parsimony-based approach to inferring gene
gain, loss, and transfer events was developed by  \citep{Doyon2011}.  We noted
that the first approach could be used to infer transfer of genomic segments due
to homologous recombination, and the second could do similarly attributed to
horizontal gene transfer.

In short, the complete genomes of the two spcies of SPY and SDE presented a
unique opportunity to study gene transfer over relatively recent time scales.
We used both of the model-based approach for inferring homologous recombination
and the parsimony-based one for detecting gene gain, loss, and horizontal
transfer. We also compared both of the results in recombining gene transfer
events to show that the parimony-based method inferred recombining gene transfer
events in similar biological functional categories to those inferred with the
model-based approach. This was the first attempt as far as we know to analyaze
SPY and SDE in genome-wide scale for detecting gene transfer events not only in
the core genome but also in the pan genome, which make the analysis complete in
terms of the amount of used genome data. Our analysis of core genomes also
showed for the first time that the net directionality of, at least, recombining
gene transfer from SPY to SDE was stronger than the reverse direction.

@@

\subsection{Recombination dirved more substitutions than did point mutation}

Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
chosen (Table S\ref{tab:genome}) to study recombining gene transfer.  The genome
alignment of the five genomes showed two distinct genome structures of SPY and
SDE (Figure S\ref{fig:mauve}). The SDE genomes were larger in length than those
of SPY. Although it was unclear whether SDD was different from either of SDE1
and SDE2 from the graphical view of the genome alignment, the spcies tree
inferred from the genome alignment showed that SDD branch was attached to the
clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The distribution of the
alignment block lengths was skewed to the right with mean of 4188 base pairs
long (Figure S\ref{fig:blocksize}), which implied that the model assumption of
geometric block length distribution might be plausible in \texttt{ClonalOrigin}
method (see section \textbf{Materials and Method} and \citealt{Didelot2010}).
The relative strength of recombination compared to mutations indicated that
reconmination-driven substitutions affected more sites than did mutation-driven
substitutions (Table S\ref{tab:clonalframe}). In the \textit{Streptococcus}
bacteria homologous recombination appeared to affect more sites than did point
mutations process.

@@

Along the genome of SDE1 population mutation rate did not appear to change much
(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
mutation rates were larger than twice the global estimate of mutation rate.  The
recombination rate appeared to change more dynamically. Espeically, the genomic
region of SDE1 near 20 kilobase pairs from the origin of replication contained
alignment blocks with bell-shaped recombination rates.  The recombinant tract
length changes in a way that we could not note any particular positions along
the genome.  There appeared to be negative correlation between recombinant tract
length and recombination rate (Figure S\ref{fig:rhologdelta}).
% \texttt{WHAT ARE THESE BLOCKS? WHAT GENES ARE LOCATED IN THE BLOCK?}
% \texttt{USE UCSC GENOME BROWSER TO FIND WHAT GENES BELONG TO THOSE BLOCKS.}

@@

Most of local gene tree topologies (see section \textbf{Materials and Methods}),
or 66.7\%, along the alignment blocks were the same as the species tree, which
allowed us to be more confident of the species tree inferred using the alignment
blocks (Table S\ref{tab:Gene-tree-topologies}).  The second most probable
topology (9.3\%) was one where SDD was the outgroup, and SPY and SDE were
sister clades. SPY and SDE appeared to share about tenth of the core genomes that
they were more closely related with each other than each to SDD's genome. The
other topologies had less than 4\% of proportions.

@@

The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Table
S\ref{tab:heatmap} for the values) shows the log base 2 ratios of the \textit{a
posteriori} number of recombinant edges for a pair of donor and recipient
species tree branches and their corresponding expected values given the species
tree and the three population parameters.  Because recombinant edges from any
species tree branch to the root branch are not possible under
\texttt{ClonalOrigin} model (see Figure S\ref{fig:clonalorigin}), the column for
the root as recipient was colored green, indicating invalid recombinant edges.
Similarly, all of the cells where a recipient branch was older than the donor,
and two branches did not share any contemporary time were colored green. The
diagonal cells were colored gray, indicating no significant number of
recombinant edges, because the observed and expected numbers of recombinant
edges were similar. As recombinant edges starting from and ending in the same
species tree branch would not change the species tree, which would not change
the likelihood much, the cells would be colored gray.  Because SDD is a strictly
veterinary pathogenic bacteria, we did not expect that SDD received much gene
transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to SDD was
-2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the recombing gene
transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD was 0.698. This
relatively higher recombining gene transfer from two SDE individuals to SDD
compared with that from two SPY individuals to SDD reflected that the SDD was
more closely related to SDE than to SPY in most of the genomes.

The genomic organization between SDD and SDE1/SDE2 were much similar (Figure
S\ref{fig:mauve}), and they formed a clade in the species tree
(S\ref{fig:tree5}). However, the gene transfer from SDD to either SDE1 or SDE2
was less frequent than the case of reverse direction; SDD served as a recipient
rather than a donor. It is not unclear why SDD received more genes from SDE1 or
SDE2 than did SDE1 or SDE2 from SDD.

Because it was expected that the recombining gene transfer within a clade would
be larger than that between clades, the recombining gene transfer log base 2
ratios between SDE1 and SDE2 (1.17 and 1.20) were about twice as large as that
from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60, and 0.39).  The
intensities of recombining gene transfer from external branches of SDE to
external branches to SPY was similar to those from SDE1 or SDE2 to SDD (0.76 and
0.70).  Either SPY1 or SPY2, and SDD appeared to be no different from the view
point of SDE1 or SDE2 since the similar values of recombining gene transfer
intensity from SDE1 or SDE2 to the other three external branches.  Surprisingly,
the recombining gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and
1.07) was comparable with those from SPY1 and SPY2 to SDE1 and SDE2 (1.25, 1.25,
1.17, and 1.29).  Effectively, the recombining gene transfer from a clade SPY1
or SPY2 to a different clade SDE1 or SDE2 seemed to be no different from that
within the clade of SPY1 and SPY2. The two external branches of SDE received
genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1 and SDE2
were in the same clade as SPY1 and SPY2.  The bacterial recombination between
SPY and SDE might happen more often than that between SDE and SDD even though
the genome configurations of SDE and SDD are more similar than each to SPY. This
might be consistent with the fact that SPY and SDE shared the same ecological
niche.  The gene transfer log base 2 ratio from SPY1 and SPY2 to SDE branch was
larger than any type of gene transfer: 1.88 and 1.94, respectively.  
Furthermore, we found a total of 58 genes for which a recombinant edge of
certain types were sampled with probability at least 0.9 by focusing on specific
types of recombinant edges.  (Table S\ref{tab:genes-transfer}).  Many of the
genes in the list were transferred from SPY to SDE rather than the reverse
direction overall.  All things considered, this clearly showed that the
recombining gene transfer was biased toward the net direction from SPY to SDE.

Figure \ref{fig:hgt-heatmap} shows the number of inferred non-recombining
horizontal gene transfer events between the different branches of the species
tree.  SDE1 and SDE2 shared the most events (26 and 34) compared to other sample
pairs.  SDD appeared to acquire many genes from the ancestral SDE branch.
Transfer events between the SPY and SDE clades were inferred in smaller numbers
(<6) compared to those within each clade (23--34). Among the clade-crossing
transfers, we did not see any significant enrichment for a specific direction,
as opposed to the result for recombining transfers from the model-based analysis
(Figure S\ref{fig:mowgli-recomb-heatmap}).

\subsection{More frequent recombining gene transfer than horizontal gene transfer}

Among the 2314 gene families the largest family consisted of 59 genes, while
nearly half of gene families (1066) had exactly one gene per species, or six
genes total (Figure S\ref{fig:famsize}).  Figure \ref{fig:Gene-duplication-loss}
shows the inferred gene counts for each extant and ancestral species. Overall,
we inferred 304 duplications, 944 recombining gene transfers, and 294
non-recombining gene transfers, comprising 23.7\% of total transfer events.  We
found that duplications made up only a small fraction of the events within the
clade (10.2\%), while horizontal transfers and losses were significantly more
frequent (41.7\% and 48.0\%, respectively). We found that duplications were most
often species-specific. Many gene losses likely represented horizontal transfers
that replaced the native gene sequence.  The species SDD and SDE were the most
expanded, both in terms of duplications and in-bound transfers.

@@

\subsection{Horizontal gene transfer is associated with virulence genes}

\texttt{Select genes with high horizontal gene transfers. Relate these to
categories of virulence genes. I should find many more virulence genes in Matt's
analysis. Use the virulence gene types from SDE2's paper. Which virulence gene
types have more gene transfer?}

We found that some virulence genes tended to have more horizontally transfers
than others. 
Streptococcal C5a peptidase (\textit{scpB}) (SDEG\_0933) was inferred to have six
horizontal gene transfer. 
Putative transposase were also found to have relatively large number of
horizontal gene transfer events.
(SpyM3\_1016, SpyM3\_1053, SpyM3\_1054)

Putative response regulator (SpyM3\_0174).
Putative maltose operon transcriptional repressor (SpyM3\_0982).
Putative histidine kinase (SpyM3\_0171).
Putative cell wall hydrolase (SpyM3\_0731, SpyM3\_0922, SpyM3\_1208).

GGS\_124 possesses 5 genes that encode putative nucleases with a secretion
signal peptide: genome-encoded streptodornase (SDEG\_0541), others,
prophage-associated deoxyribonuclease (SDEG\_1103), all of which are predicted to code for a secretion signal peptide.
Streptodornase (Sdn) - phage associated  (SpyM3\_1409).

Protein F2 like fibronectin-binding protein (SpyM3\_0104).

\subsection{Comparison between model-based and parsimony-based approaches}

\texttt{I might want to analyze only core genomes and pan genomes
separately. By separating data set of 6-gene families from the other families,
we could have more complete comparison.}

\texttt{Genes and recombining gene transfer events count and recombination
intensity might be use to compare two methods.}

@@

\section{Discussions}

Closely related bacterial species were investigated here to better understand
gene transfer processes. Two complementary methods were employed to study
recombining and non-recombining gene transfers. By discriminating the two
processes in analyzing the genomes, we provided evidence that the pathogenicity
of SDE and SPY was more attributed to non-recombining gene transfer process than
homologous recombining gene transfer.  Direction of gene flow between SDE and
SPY has been questioned and challenged
\citep{Kalia2001,Kalia2004,Towers2004,Bessen2005,Davies2007,Bessen2010} in order
to elucidate the origin of opportunistic human pathogenicity of SDE via transfer
from SPY into SDE.  We also showed in this study that the amount of recombining
gene flows from SPY to SDE was relatively greater than that of the reverse
direction using the core genome sequences.

\textbf{Virulence genes and gene transfer:} 
Because we found directional recombining gene transfer from SPY to SDE, and did
not found virulence genes association with recombining gene transfer, this
indicated that pathogenicity of SDE might not be due to acquisition of virulence
genes from SPY via recombining transfer. This does not necessarily mean that
virulence genes were not exchanged between SPY and SDE via non-recombining
transfer.  Because the \texttt{Mowgli} analysis just did not reveal unbalanced
non-recombining or recombinding gene transfer, we suspected that \texttt{Mowgli}'s
parsimony might not be powerful in detecting unbalanced gene flow.  We
conjecture the following hypothesis of the evolution of SPY and SDE to support
the non-recombining virulence gene exchange between SPY and SDE.  Pathogenicity
of SDE might have existed before the common ancestor of SPY and SDE.  We
speculated that SPY might have been split off from the common ancestor that was
SDE-like because the population size of SDE was larger than that of SPY (CITE).
After the split, SPY might have been specialized in inhabting human hosts, and
SDE in colonizing veterinary hosts. Recent surge in cases of human diseases
attributed to SDE might be because of secondary contact of SPY and SDE since
their split from their common ancestor.  When SPY thrives in its environment of
human hosts and another relatively close species, SDE, starts to cohabit with
the indigenous species SPY, genetic elements in one of the two species could
benefit another species. If the two species do not compete with each other, then
the increased genetic variation from multiple species would eventually bring
advantage to both of the species.  Virulence genes could have been beneficial to
both species in inhabting human hosts.  Because the common ancestry of SPY and
SDE, recombining gene transfer between the two species might be relatively
prevalent. But, these homologous recombination might have rarely contributed to
influx of virulence genes from SPY to the opportunistic human pathogen of SDE.
Because homologous parts of SPY and SDE would be similar in their genetic
compositions, the homologous parts would be unlikely to contain virulence genes
to human hosts. We suspected that virulence genes were exchanged between SPY and
SDE, but did not find any directionality of the gene flow.

@@

\section{Supplementary Information}

\subsection{Population mutation and recombination rates were accurately
estimated}

% FIXME: Newick tree in text in a file.
We describe the detailed procedure of simulations for checking if the three
population parameters were accurately recovered from the first stage of
\texttt{ClonalOrigin} MCMC.  Ten recombinant trees were sampled given the three
parameters and the species tree  set to the values inferred from the five
bacterial genomes (see section \textbf{Results}).  Each of the recombinant trees
was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
274 blocks of alignments.  The block lengths configuration was the same as that
of the blocks used in the analysis of the five genomes.  The mean of the ten
estimates of the recombinant tract length was significantly higher than the true
value of (919 bp versus 744 bp) whereas the values of the mutation and
recombination rates were accurately estimated (Table
S\ref{tab:sim-three-population-parameters}).  Recombinant tract lengths were
known to be difficult to estimate from relatively short alignment blocks
\citep{Didelot2010}. Performing similar simulations with much longer blocks
(i.e., 10000 bp) we estimated the average recombinant tract length accurately
\texttt{(Table S data not shown)}.

\subsection{Estimating numbers of recombinant edges were accurate}
